---
title: "Python environments"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Python environments}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

skald uses `reticulate::py_require()` to declare the Python requirements for
PyLate. This keeps package loading lightweight and lets reticulate resolve a
compatible Python environment when needed.

## Default requirements

```{r}
library(skald)

skald_configure()
```

The default requirement is `pylate>=1.4.0,<1.5` with Python 3.10 or newer.

## Optional extras

Use PyLate extras when a workflow needs them:

```{r}
skald_configure(extras = "eval")
skald_configure(extras = "voyager")
```

## Diagnostics

`skald_diagnostics()` reports R, reticulate, Python, PyLate, torch, transformer,
and device information.

```{r}
skald_diagnostics()
skald_torch_devices()
skald_check_install()
```

Call diagnostics before debugging model downloads or GPU behavior.
