---
title: "Quick start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

skald is an R interface to late-interaction retrieval. It wraps Python PyLate
objects while exposing tidy R workflows for candidate reranking, local indexing,
and retrieval.

## Configure Python

Call `skald_configure()` before Python is initialized if you need to control
the active PyLate requirements.

```{r}
library(skald)

skald_configure()
skald_diagnostics()
```

## Create a model

```{r}
model <- skald_model("lightonai/GTE-ModernColBERT-v1", device = "cpu")
```

The returned `SkaldModel` stores enough R-side metadata to be printed,
serialized, and rehydrated later. The Python model object stays opaque.

## Rerank candidates

```{r}
candidates <- tibble::tibble(
  doc_id = c("a", "b", "c"),
  text = c(
    "dplyr::filter() keeps rows that match logical conditions.",
    "dplyr::select() keeps columns.",
    "mutate() creates or changes columns."
  )
)

skald_rerank(
  candidates,
  model,
  query = "How do I subset rows?",
  text = text,
  id = doc_id,
  top_k = 2
)
```

`skald_rerank()` preserves input columns, appends `late_score` and `late_rank`,
and sorts by late-interaction score.
