---
title: "Rerank ragnar results"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rerank ragnar results}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

ragnar is the natural first-stage retriever for many RAG workflows. skald is
designed to sit after ragnar, reranking the candidate rows with a late-
interaction model.

```{r}
library(ragnar)
library(skald)

model <- skald_model("lightonai/GTE-ModernColBERT-v1", device = "cpu")

query <- "How can I join two tibbles?"

hits <- ragnar_retrieve(store, query, top_k = 50)

hits |>
  skald_rerank(
    model,
    query = query,
    text = text,
    id = c(origin, doc_id, chunk_id),
    top_k = 8
  )
```

`ragnar_retrieve()` names its second argument `text`, so pass the query string
positionally. `skald_retrieve_ragnar()` handles that detail for one-call
retrieval plus reranking:

```{r}
skald_retrieve_ragnar(
  store,
  model,
  query = "How do I use dplyr filter?",
  candidate_k = 40,
  top_k = 6,
  filter = area == "data-transform"
)
```

The returned tibble keeps ragnar columns and adds `late_score` and `late_rank`.
