---
title: "Skald stores"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Skald stores}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

`SkaldStore` combines a DuckDB metadata sidecar with a PyLate PLAID index.
This gives R users a retrievable local store without treating ColBERT
multi-vector embeddings as ordinary dense vectors.

## Create and insert

```{r}
library(skald)

model <- skald_model("lightonai/GTE-ModernColBERT-v1", device = "cpu")

store <- skald_store_create(
  "docs.skald",
  model = model,
  name = "docs",
  title = "Documentation"
)

chunks <- tibble::tibble(
  origin = "example.md",
  start = c(1L, 80L),
  end = c(79L, 140L),
  context = c("Filtering rows", "Selecting columns"),
  text = c(
    "Use filter() with a logical expression to keep rows.",
    "Use select() to keep or reorder columns."
  )
)

skald_store_insert(store, chunks)
```

For chunks with `origin`, `start`, and `end`, skald creates stable string IDs
from the source and character positions.

## Build and retrieve

```{r}
skald_store_build_index(store)

skald_retrieve(store, "How do I keep rows?", top_k = 5)
```

`late_score` is a MaxSim score. Higher values are better within a query, but
the score is not a probability and is not suitable for universal thresholds.
