---
title: "Token highlighting"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Token highlighting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Late-interaction models score a document by matching each query token to its
best document token. `skald_token_relevance()` exposes those token-level
matches, and `skald_token_highlight()` renders them as HTML.

## Compute relevance

This is the live workflow against the R4DS data transformation chapter:

```{r, eval = FALSE}
library(dplyr)
library(ragnar)
library(skald)

model <- skald_model("lightonai/GTE-ModernColBERT-v1", device = "cpu")

query <- "How do I keep rows with filter?"

r4ds_chunks <- "https://r4ds.hadley.nz/data-transform.html" |>
  ragnar::read_as_markdown() |>
  ragnar::markdown_chunk() |>
  tibble::as_tibble()

candidates <- r4ds_chunks |>
  filter(grepl("filter", text, ignore.case = TRUE)) |>
  slice_head(n = 3)

token_scores <- skald_token_relevance(
  model,
  query = query,
  documents = candidates$text,
  document_id = seq_len(nrow(candidates))
)
```

`token_scores` is a tibble with one row per document token. The most important
columns are:

- `score`: raw best query-token similarity for the document token.
- `score_scaled`: display scale in `[0, 1]`.
- `best_query_token`: query token that best matched this document token.
- `selected_by_query`: whether at least one query token selected this document
  token under MaxSim.

## Render highlights

The package site uses a small deterministic R4DS-style excerpt so the article
can render without downloading a model. The object has the same columns returned
by `skald_token_relevance()`.

```{r}
library(skald)

r4ds_excerpt <- paste(
  "The R4DS data transformation chapter introduces filter() as a way to keep",
  "rows that satisfy a logical condition. This makes it useful when a question",
  "asks how to subset observations before summarising or plotting data."
)

r4ds_token_scores <- tibble::tribble(
  ~token,          ~score, ~score_scaled, ~best_query_token, ~selected_by_query,
  "filter()",       0.92,          1.00,  "filter",          TRUE,
  "keep",           0.78,          0.78,  "keep",            TRUE,
  "rows",           0.73,          0.69,  "rows",            TRUE,
  "logical",        0.61,          0.47,  "with",            FALSE,
  "condition",      0.66,          0.56,  "filter",          FALSE,
  "subset",         0.84,          0.86,  "keep",            TRUE,
  "observations",   0.57,          0.40,  "rows",            FALSE
)

loc <- vapply(
  r4ds_token_scores$token,
  function(token) regexpr(token, r4ds_excerpt, fixed = TRUE)[[1]],
  integer(1)
)

r4ds_token_scores <- dplyr::mutate(
  r4ds_token_scores,
  document_id = "r4ds-data-transform",
  document = r4ds_excerpt,
  token_start = loc,
  token_end = loc + nchar(.data$token) - 1L,
  .before = 1
)

htmltools::tags$div(
  style = "max-width: 72ch; font-size: 1.05rem; line-height: 1.8;",
  skald_token_highlight(r4ds_token_scores)
)
```

Hover over a highlighted token to see its raw score and best matching query
token.

```{r}
r4ds_token_scores |>
  dplyr::select(token, score, score_scaled, best_query_token, selected_by_query)
```
